---
title: "Getting Started with semanticfa"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with semanticfa}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5
)
```

## Overview

`semanticfa` performs exploratory factor analysis on language model embeddings
of psychological scale items. Given item text, it embeds each item, computes a
similarity matrix, and extracts latent factors — entirely from the text, with
no human response data required.

The package is designed to feel familiar to `psych` and `EFAtools` users.

## Quick start with bundled data

The package ships with the 50-item IPIP Big Five inventory and precomputed
Qwen3-Embedding-8B embeddings (50 x 4096, rounded to 4 decimal places), so you
can try it with zero setup:

```{r quickstart}
library(semanticfa)
data(big5)

fit <- sfa(
  big5$items,
  nfactors    = 5,
  embeddings  = big5$embeddings,
  scoring     = big5$scoring
)
print(fit)
```

## Factor retention

When you omit `nfactors`, `sfa()` uses embedding-adapted parallel analysis
(random unit vectors in the embedding dimension as the null):
```{r retention}
fit_auto <- sfa(
  big5$items,
  embeddings = big5$embeddings,
  scoring    = big5$scoring
)
cat("Auto-detected factors:", fit_auto$factors, "\n")
```

For a multi-method comparison, use `sfa_nfactors()`:
```{r nfactors}
sim <- sfa_similarity(big5$embeddings, encoding = "atomic_reversed",
                      scoring = big5$scoring)
nf <- sfa_nfactors(sim, big5$embeddings,
                   methods = c("parallel", "kaiser"),
                   parallel_iter = 50)
print(nf)
```

## Encoding methods

The `encoding` argument controls how embeddings become a similarity matrix:

```{r encoding}
sim_ar  <- sfa_similarity(big5$embeddings, "atomic_reversed", big5$scoring)
sim_sq  <- sfa_similarity(big5$embeddings, "squid", big5$scoring)
sim_mcp <- sfa_similarity(big5$embeddings, "mean_centered_pearson", big5$scoring)

cat("atomic_reversed range:", range(sim_ar[lower.tri(sim_ar)]), "\n")
cat("squid range:          ", range(sim_sq[lower.tri(sim_sq)]), "\n")
cat("mean_centered_pearson:", range(sim_mcp[lower.tri(sim_mcp)]), "\n")
```

The ranges show what each encoding does to the sign structure. Sign-flipping
(`atomic_reversed`) manufactures strong negatives by turning reverse-keyed
items into anti-topic vectors, and SQuID's questionnaire-mean centering
recovers a modest negative range. Mean-centered Pearson tracks the raw
cosines, which stay positive for these items.

## Visualization

```{r scree, fig.cap="Scree plot with parallel analysis threshold"}
plot(fit, type = "scree")
```

```{r loadings, fig.cap="Factor loading heatmap"}
plot(fit, type = "loadings")
```

## Comparing with empirical factor analysis

The `$loadings` component works directly with `psych` functions:

```{r psych-compat, eval=FALSE}
# Run human-data EFA (not run — requires response data)
human_fit <- psych::fa(response_data, nfactors = 5, rotate = "oblimin")

# Compare
psych::factor.congruence(fit$loadings, human_fit$loadings)
```

For NMI, ARI, Frobenius, and disattenuated correlation:

```{r congruence}
cong <- sfa_congruence(fit, big5$factors, metrics = c("nmi", "ari"))
print(cong)
```

## Using your own embeddings

Pass any embedding model's output via `embeddings=`:

```{r custom, eval=FALSE}
# With sentence-transformers (requires reticulate + Python).
# The default model is "Qwen/Qwen3-Embedding-0.6B"; larger models such as
# "Qwen/Qwen3-Embedding-4B" (8 GB RAM) or "Qwen/Qwen3-Embedding-8B" (16 GB RAM)
# recover factor structure more accurately.
emb <- sfa_embed(my_items, embed = "sbert", model = "Qwen/Qwen3-Embedding-0.6B")
fit <- sfa(my_items, embeddings = emb, scoring = my_scoring)

# Or bring your own function
my_embedder <- function(texts) {
  # ... your embedding logic ...
  # must return a numeric matrix (n_items x dim)
}
fit <- sfa(my_items, embed = my_embedder, scoring = my_scoring)
```
