% sections/abstract.tex  (input in the preamble, before \begin{document})
\abstract{%
The items of a psychometric scale constitute a proposal, written in ordinary
language, about the structure of a construct---and language-model embeddings
make that proposal measurable before a single respondent is recruited. We
introduce \texttt{semanticfa}, an R package for semantic factor analysis
(SFA): exploratory factor analysis of the similarity structure of
language-model embeddings of scale items, requiring no response data at any
stage. The package implements the complete workflow: similarity construction
under keyed, keying-free, and natural-language-inference encodings;
factor retention by multi-criterion consensus; extraction with Monte
Carlo--calibrated diagnostics; interpretation by anchoring, semantic
projection, congruence, and jingle--jangle screening; and response-free
refinement (redundancy detection, short forms, candidate-item vetting). A
fully reproducible demonstration
exercises every exported function on the 50-item International Personality
Item Pool Big-Five Factor Markers, embedding the items with
Qwen3-Embedding-8B and validating the semantic structure against a
conventional factor analysis of 874,434 respondents. The retention consensus
indicated five factors, and the semantic solution recovered the Big Five
domains, with congruence to the human factors graded by domain (Tucker's
$\phi$ from .92 for Openness to .52 for Agreeableness) and semantic
similarities correlating $r = .46$ with response correlations across all
1,225 item pairs. Sign-flipping reverse-keyed embeddings produced anti-topic
vectors that aligned worst with keyed human data, and a semantically central
25-item short form recovered the theoretical structure better than the full
inventory. We frame the agreement of the two structures---semantic--behavioral
coherence---as an estimable relation in a construct's nomological network: a
complement to, not a replacement for, response-based factor analysis. The
package is available on CRAN.
}
\keywords{semantic factor analysis, language models, embeddings, construct validity, scale development, R package}
