% discussion_coherence.tex
\subsection{Theoretical Implications}

The demonstration shows what estimating the Introduction's
semantic--behavioral coherence buys. In the terms of
\citet{cronbach-meehl-1955}, the semantic and human analyses are
qualitatively different operations (item-meaning geometry without
respondents versus \valNHuman{} people's response covariance), so their
near-match for Openness, Conscientiousness, and Neuroticism ($\phi =
\valPhiO$, \valPhiC, and \valPhiN) is intranetwork evidence that both engage the same
network region. Under the realist caveat conceded at the outset, this
strengthens the evidential webs without certifying that the attributes
exist or cause anything \citep{borsboom-etal-2004}: for these domains,
the operationalized semantic proposal and response structure are two
views of one organization.

The divergences (localized and interpretable, not diffuse) earn the
relation its keep: the semantic solution splits Agreeableness and
Extraversion into warmth/other-orientation and gregariousness factors
rather than losing them, semantic Extraversion thereby matching human
Agreeableness ($\phi = \valPhiEA$) nearly as well as semantic Agreeableness did
(Table~\ref{tab:tucker}). This is a seam where English usage runs sociability
and warmth together while response covariance holds them apart. The
pattern has precedent: \citet{kojima-etal-2026} found
near-interchangeability at the general psychopathology factor (congruence
$= .97$) but specific-factor divergences of non-semantic origin (insomnia
fused with somatic complaints only in responses, plausibly
physiological). \citet{kmetty-etal-2021} found health-related occupations
lower in embedding-derived prestige than in survey standing, plausibly
via illness connotations, and \citet{zhang-qiu-2026} found the relation
construct-dependent, weakest where baseline inter-item cohesion is
strongest. Divergence thus signals its source: semantic compression of
distinct behavioral dimensions, connotational drift, or behavioral
covariance with causes wording does not carry.

Making coherence estimable also turns the Introduction's divergence
conditions into testable predictions. If, as \citet{borsboom-2008}
argues, some clinical constructs are causal symptom networks, coherence
should run systematically lower there than for common-cause
constructs, an ordering \texttt{sfa\_congruence()} makes falsifiable
across instruments of both kinds. The logic extends to formative
indicator sets \citep{bollen-lennox-1991} and method variance
\citep{messick-1995}. An unexamined confound of every rating-scale factor
analysis becomes a quantity theories can be required to predict.
