% discussion_limitations.tex
\subsection{Limitations}

The central limitation is definitional: semantic factor analysis
characterizes wording, not people. With no respondent, it asserts no
causal path and supplies none of the response-process theory
\citet{borsboom-etal-2004} make the heart of validity (absent even from
response-based factor structure), so coherence must never be reported as
a validity coefficient. Of the six construct-validity aspects in
\citet{messick-1995}, semantic evidence informs content (domain
boundaries, representation, redundancy) and partly structural (scoring
mirroring the semantic proposal). The substantive, generalizability,
external, and consequential aspects require response and outcome data. Even the
coherence thesis's strongest support bounds it: in
\citet{kojima-etal-2026}, response-based factor scores predicted lifetime
diagnoses better than text-based ones, textual information being
effectively a subset of the behavioral. Hence the calibrated framings: a
free, instant pilot study always followed by human data
\citep{hommel-arslan-2025}, no edge over aggregated expert judgment
\citep{schoenegger-etal-2025}, and, in \citet{feraco-toffalini-2025}, no
false promises of empirical fit though roughly half of misfit verdicts
were false alarms (confidence better grounded than condemnation).

High coherence prices rather than rebuts the critique of
\citet{uher-2025}, simultaneously grounding pre-data screening and
measuring the share of ``nomological'' structure from ratings already in
the items' inbuilt semantics, the quantity the semantic theory of survey
response first estimated with LSA-era tools
\citep{arnulf-etal-2014, arnulf-etal-2018}: here $r = \valPairR$ at the
pair level, $\bar{\phi} = \valPhiMean$ for factor shapes. Meaning-as-use also bounds the
method: embeddings summarize a community's conventionalized regularities
of use \citep[\S43]{wittgenstein-1953}, not the local construals raters
form, so the idiosyncratic share of response variance is missed by design
and perfect coherence is unexpected even from a flawless instrument.

A further limitation is specific to the demonstration corpus. The IPIP
Big-Five Factor Markers are among the most widely reproduced psychometric
texts on the public web, frequently listed together with their domain
labels, including in the codebook of the very benchmark used here
\citep{openpsychometrics-2018}, and the pretraining corpus of the
embedding model is not public \citep{qwen-2025}. Embedding proximity for
these items may therefore partly encode documented scale co-membership
rather than the compositional meaning of the sentences, which would
inflate agreement with the human structure. The reorganization of
Agreeableness and Extraversion (Table~\ref{tab:tucker}) argues against
pure memorization of the published key, but contamination cannot be
excluded without deliberately reworded materials, so coherence estimated
on canonical inventories should be read as an upper bound. The package's
central use case, vetting new and unpublished items, is precisely the
case a memorized taxonomy cannot cover, and paraphrase-based robustness
probes (re-embedding systematically reworded items) are a natural
response-free check left to future work.

The Monte Carlo machinery carries a related caveat. The
parallel-analysis and calibration nulls use random unit vectors, which
are nearly orthogonal in high dimension, so the null similarity matrix is
close to an identity matrix: its eigenvalue thresholds concentrate just
above one, and beating a structureless baseline is a weak test for
matrices whose defining feature is a large shared positive component.
In practice this makes the adapted parallel analysis operate as a
slightly elevated latent-root criterion: its verdicts track eigenvalue
magnitude against a threshold near one rather than testing where
interpretable structure ends, and we flag repairing this null as an
open methodological problem (see Future Directions), not a settled
feature.
The corpus-resampled null of \citet{pokropek-2026} retains the baseline
thematic similarity that real items share, and the one benchmark of
eigenvalue methods on embedding similarity matrices found parallel
analysis systematically overextracting \citep{garrido-etal-2025}.
The comparison-data analysis in the Supplementary Materials documents
the difficulty from the model side: on these embedding matrices,
reproduction of the observed spectrum keeps improving smoothly with
every added comparison factor, so the sequential stopping rule of
\citet{ruscio-roche-2012} saturates at its cap, whereas the same
procedure applied to the human responses shows a sharp elbow at exactly
five factors. The spectrum of an embedding similarity matrix appears to
carry graded structure with no crisp factor boundary, so single-count
verdicts on it are depth choices rather than discoveries.
Retention conclusions here therefore lean on the multi-criterion
consensus rather than on any single rule, and reference distributions
matched on the observed general similarity component are the needed
refinement (see Future Directions). The human benchmark is likewise a
convenience sample: an open, self-selected internet administration
\citep{openpsychometrics-2018} cleaned only by response-range screening,
so its factor structure need not match that of a probability sample, and
the substitutability bound of \citet{guenole-etal-2025}, the match
between the embedding model's training population and the respondent
population, applies to this comparison as to any other.

Measurement is also model-dependent: corpus register, training window,
and architecture change which human behaviors embedding similarities
explain, none optimal across tasks
\citep{mandera-etal-2017}. The encoder is thus a declared instrument
whose robustness merits reporting. The results rest on one model family
(Qwen3-Embedding) and English items. Even default full vectors can be
suboptimal, signal concentrating in coordinate subsets
\citep{golino-2026}: witness the optimal depth of \valDepthOpt{} of \valDepthFull
here. Embedding methods trail human validators on behaviorally but not
lexically phrased items \citep{milano-etal-2025b}, so behavioral
instruments may recover less well than the trait-adjacent Big-Five Factor
Markers. Psychopathological nuance and hard-to-verbalize inner
experience, likely underrepresented in training corpora, leave semantic
structure least trustworthy where clinical assessment is hardest
\citep{kambeitz-etal-2025}.

Finally, the refinement tools carry caveats. Pruning presupposes
reflective indicators. Formative indicators are a census, not a sample,
and removing one changes the construct \citep{bollen-lennox-1991}, so
the measurement model must be settled first. The UVA cutoffs were
calibrated on simulated response data, not semantic similarity matrices
\citep{christensen-etal-2023}. Semantically close items can still differ
in response process (difficulty, extremity, framing) or in theoretical
identity, as with subjective stress versus the cortisol stress response.
Similarity is not measurement equivalence \citep{wang-etal-2026}.
Every purely semantic redundancy or
jingle--jangle verdict (this package's included) should, as
\citet{wulff-mata-2026} argue, inform expert deliberation, never replace
it.
