% introduction_prediction.tex
\subsection{Embeddings Predict Factor Structure and Item Correlations}

A growing literature shows embedding similarity recovers inventories' a-priori
factor structure without responses. \citet{milano-etal-2025a} embedded four scales
(Big-Five markers, RIASEC, Humor Styles, DASS), and a principal component analysis
of cosines loaded over 70\% of items on their predicted construct, loadings
correlating with response loadings; \citet{casella-etal-2024} report the same for
the Big Five. \citet{guenole-etal-2025} formalize a \emph{substitutability
assumption}---an item's embedding stands in for its response vector---and run
\emph{pseudo-factor analysis} on the cosine matrix, recovering the five- and
six-factor IPIP NEO-300 and HEXACO-240 structures.

The convergence comes with informative divergence, not equivalence:
\citet{guenole-etal-2025} find Tucker congruence often below conventional
thresholds even as matched factors correlate and non-matched do not, while
\citet{milano-etal-2025a} and \citet{casella-etal-2024} report agreeableness and
extraversion items bleeding together, mirroring their elevated behavioral
correlation---semantic compression of distinct dimensions. \citet{guenole-etal-2025}
bound substitutability by training--response population similarity and intend
pseudo-factor analysis as a pre-trial check---the response-free use case
\citet{milano-etal-2025a} and \citet{milano-etal-2025b} also state.

The most direct evidence predicts the empirical inter-item correlations.
\citet{hommel-arslan-2025} fine-tuned a sentence transformer so cosine proximity
estimates behavioral correlation, recovering out of sample, across 79 scales,
interitem correlations at $r=.59$ and scale intercorrelations and reliabilities at
$r=.84$. To overcome polarity---cosines rarely turn negative---they relabeled the
SNLI corpus with signed magnitudes, training the model to return negative
similarities for contradiction, letting a similarity matrix stand in for a
covariance matrix \citep{hommel-arslan-2025, bowman-etal-2015}.
This is not confined to fine-tuned encoders: \citet{ravenda-etal-2025} found an
item's strongest empirical partner among its three nearest semantic neighbors in
82\% of Big-Five and 95\% of DASS cases, an EFA reproducing the response structure.
It even rivals expertise: \citet{schoenegger-etal-2025} found AI outpredicted most
individual humans, experts included, on item-pair correlations, though the residual
gap cautions these approximate rather than replace empirical values. It is also
partial: fitting confirmatory models to the cosine matrix against 31,697 VIA-IS-P
respondents, \citet{feraco-toffalini-2025} found semantic and empirical correlations
agreeing only moderately ($r=.67$)---every semantically fitting model fit the data,
but only half of misfits truly misfit---so coherence is evidence, not identity, and
SFA a complement to, not a substitute for, data collection.
