% introduction_package.tex
\subsection{The Present Package}

These threads converge on one opportunity. Item geometry encodes relational
knowledge---ordering items between antonymic anchors recovers human feature
judgements from the embedding alone, though intra-linguistically
\citep{grand-etal-2022}; embeddings thus capture the semantic geometry of a scale's
language \citep{westbury-preprint}, which for our purposes is its \emph{wording}
rather than the attribute itself. What has been missing
is a single, general-purpose, response-free toolkit bringing the full discipline of
exploratory factor analysis to that structure in one reproducible workflow.

\texttt{semanticfa} \citep{yanitski-westbury-2026} is that toolkit. Built on R
\citep{rcoreteam-2025} and interoperating with the factor-analytic and
network-psychometric ecosystem \citep{revelle-2026, golino-christensen-2026}, it
implements SFA in five stages: \emph{similarity construction} under keyed,
keying-free, and natural-language-inference encodings; \emph{retention} by
multi-criterion consensus; \emph{extraction} with Monte Carlo--calibrated
diagnostics; \emph{interpretation} by anchoring, semantic projection, congruence,
and jingle--jangle screening; and response-free \emph{refinement}. Semantic and
behavioral solutions are compared by the chance-corrected adjusted Rand index
\citep{hubert-arabie-1985}.

Positioning clarifies what is new. The name \emph{semantic factor analysis} was
coined independently by \citet{muller-2026} for a Word2Vec recovery of Big-Five
structure from adjectives \citep{mikolov-etal-2013, goldberg-1990}, each scaled by
its mean response and partitioned. Our use denotes a distinct, broader method: a
fully response-free factor analysis of the embedding similarity structure, not
restricted to adjectives---answering, rather than committing, the cardinal error
\citet{uher-2025} identifies.

We demonstrate every exported function on the 50-item IPIP Big-Five Factor Markers,
embedded with Qwen3-Embedding-8B \citep{qwen-2025} and validated against a
conventional factor analysis of \valNHuman{} Open-Source Psychometrics respondents
\citep{openpsychometrics-2018}, asking whether consensus retention recovers the Big
Five. We turn now to the method.
