% introduction_pipelines.tex
\subsection{Methods, Pipelines, and the Wider Applied Wave}

A first cluster treats the encoding step. Because embeddings compress similarities
toward the positive---so that pseudo-factor loadings do not differentiate
reverse-keyed items \citep{suarez-alvarez-etal-2026}---SQuID subtracts the mean
``questionnaire embedding'' from each item, recovering a signed range without
fine-tuning \citep{pellert-etal-2026}. Fitting confirmatory
models to mean-centered cosines, \citet{pokropek-2026} shows via Monte Carlo that
fit indices stay informative with six or more indicators but are met by chance with
few. And Unique Variable Analysis detects local dependence via a regularized network
and weighted topological overlap \citep{christensen-etal-2023}, a screen that
transfers to the semantic matrix because shared semantic reference drives such
dependence.

A parallel line transports network-psychometric retention onto the embedding matrix.
Exploratory Graph Analysis (EGA) estimates dimensionality by community detection
\citep{golino-2026, garrido-etal-2025}: benchmarking six facets,
\citet{garrido-etal-2025} found Kaiser's rule \citep{kaiser-1960} and parallel
analysis \citep{horn-1965} overestimated dimensionality while EGA recovered it,
near-perfectly after weighted-topological-overlap filtering
\citep{christensen-etal-2023}; \citet{golino-2026} adds Dynamic EGA to optimize
recovery.
The geometry also supports response-free screening: \citet{jung-seo-2025}
selected centroid-proximal IPIP markers by $k$-means, recovering a 30-item form
matching response-based reductions \citep{openpsychometrics-2018}; and
\citet{wang-etal-2026} clustered contextual embeddings by density, cutting items by
$\approx60\%$ across DASS, IPIP, and EPOCH while preserving structure---both relying
on chance-corrected partition agreement \citep{hubert-arabie-1985} and projection
\citep{vandermaaten-hinton-2008}. \citet{hussain-etal-2024} run on-device models,
showing item cosines track response correlations; and \citet{low-etal-preprint},
flagging a ``psychometric bias'' judging models only by label prediction, name
embedding-based factor analysis as open.

A generative wave drafts items rather than dissecting them: the AI-GENIE framework
\citep{russell-lasalandra-etal-2026} has a model write a pool pruned in silico
(EGA, UVA, bootstrap EGA), reported to match representative samples for the
best-performing models. Others dissect existing items: \citet{huang-etal-2025}
cluster item embeddings to flag redundancy, recovering in a preliminary study
Grit--Conscientiousness overlap while keeping gratitude separate. A parallel strand
uses models as respondents: \citet{liu-etal-2025} calibrate LLM responses with item
response theory, obtaining difficulties correlating with human values but narrow
abilities---still a person-by-item matrix. Models also stand in at construction:
\citet{eberhardt-etal-2025} build an
\emph{LLM rating scale} from a local model's transcript ratings, reaching strong
reliability; yet \citet{grobelny-etal-2025} found a ChatGPT-4 adaptation matching
the human one on structure but with lower internal consistency. Validation is then
constitutive: \citet{bunt-etal-2025} show a model's text coding earns trust only
after iterative semantic, predictive, and content validity---an \emph{abductive}
use. \citet{lim-etal-2025} screen items with
simulated respondents on the Big Five, Schwartz values, and VIA; and
\citet{maharjan-etal-2025} scrutinized embeddings as trait measures across
$\sim$1M Reddit posts. Urgency comes from a threat to response data:
\citet{westwood-2025} showed an autonomous agent passing $99.8\%$ of attention
checks, so text-only checks are immune to synthetic-respondent fraud.

A further line asks whether item language carries item-response-theory parameters.
Synthesizing 37 articles, \citet{peters-etal-2025} find text recovers
\emph{difficulty} usefully ($r$ up to $.87$); \citet{han-etal-2025} find
\emph{discrimination} largely text-resistant; and \citet{ormerod-2026} recover it
better by simulating graded-ability responses with fine-tuned generative models. The case is sharpened by critique: \citet{uher-2025}
argues statistics is not measurement and warns of a \emph{cardinal error}---mistaking
judgments of verbal statements for measurements of what they describe; yet this same
critique supplies the divergence mechanism, items' inbuilt semantics accounting for
much rated covariance. Finally, the tools carry their provenance: the
sentence-similarity backbone derives from natural-language-inference training on the
SNLI corpus \citep{bowman-etal-2015}, which first enabled cosine-comparable
embeddings \citep{reimers-gurevych-2019}; visualization uses $t$-SNE
\citep{vandermaaten-hinton-2008}; and interrogation uses semantic projection along
antonymic anchors \citep{grand-etal-2022}.
