% introduction_taxonomy.tex
\subsection{Taxonomy, Jingle--Jangle, and Clinical Structure}

Embedding analyses also address \emph{taxonomic incommensurability}, the difficulty
of comparing constructs across traditions that map theory onto measurement
incompatibly \citep{wulff-mata-2025}. \citet{wulff-mata-2025} placed 4,452 IPIP
items, 459 scales, and 277 labels in a shared space, recovering known psychometric
properties, and used the joint content--label embedding to automate
\emph{jingle--jangle} detection (a jingle when same-labeled scales differ
semantically, a jangle when similar scales carry distinct labels), flagging 504
fallacies and a taxonomy with $\approx75\%$ fewer labels.
\citet{wulff-mata-2026} generalize this as scalable \emph{conceptual
engineering}, insisting it inform rather than replace expert judgment, since
embeddings inherit training biases. Automated screening of this kind has
LSA-era precedents: \citet{larsen-bong-2016} detected construct-identity
fallacies across published scales with language-processing designs
competitive with doctoral-student searchers, outperforming them on less
common constructs, and the Semantic Scale Network of
\citet{rosenbusch-etal-2020} quantifies a new scale's semantic overlap with a
scale corpus, supporting redundancy review before any data collection.

The logic extends to clinical items, themselves verbal symptom descriptions.
\citet{kambeitz-etal-2025} embedded four psychopathology scales (DASS, O-LIFE, AQ,
Prodromal Questionnaire), found cosines tracking empirical item-pair correlations,
and showed $k$-means clustering reproduced the established subdomains, the embedding
and empirical partitions agreeing above chance by the adjusted Rand index
\citep{hubert-arabie-1985}: better than chance, worse than the data. The contrast is sharpest against an external criterion: across 145 symptom
items, \citet{kojima-etal-2026} extracted factors from item-embedding cosines and
from
response polychorics, finding the general psychopathology factor nearly
interchangeable (congruence $=.97$) yet specific factors diverging of non-semantic
origin. Insomnia and somatic items fused only in responses, the syndromic factor
predicting diagnostic histories better than the semantic one. On the same corpus we
revisit \citep{openpsychometrics-2018}, \citet{stanghellini-etal-2024} introduced
\emph{semantic loadings} (overlaps between cognitive-network communities of item
concepts and exploratory-graph factors of responses), resolving emotional from
physical anxiety facets EFA left merged.

The relation is construct-dependent: modeling 463 item pairs,
\citet{zhang-qiu-2026} found cosine similarity predicted inter-item correlations
but with large random-slope variance and a negative intercept--slope
correlation. Where baseline cohesion is high, similarity adds less. The same logic
recovers social structure:
\citet{kmetty-etal-2021} factor-analyzed a similarity matrix over 234 occupation
terms, extracting dimensions correlating $r\approx.71$--$.78$ with the survey-based
socio-economic index, with informative divergences (health-care occupations
under-ranked, an organizational-power dimension emerging from text alone). That word
meaning is itself a measurable resource is long established: \citet{buchanan-etal-2001}
showed semantic-neighborhood measures from a co-occurrence space predict word
recognition, with a precisely quantifiable advantage over feature-based
measures.\looseness=-1
