% open_practices.tex — Behavior Research Methods requires an Open Practices
% Statement immediately prior to the References section. Content moved from
% the former Method subsection (methods_open_practices.tex).
\section{Open Practices Statement}

A self-contained reproduction archive, hosted on the Open Science Framework
at \url{https://osf.io/XXXXX}, regenerates every number, table, and figure in
the demonstration from raw inputs. It contains the master script
(\texttt{reproduce.R}) and the jackknife interval script
(\texttt{make\_ci\_assets.R}), a full-run console transcript (\texttt{reproduce.Rout})
for checking every command against its real output, the precomputed
Qwen3-Embedding-8B item and auxiliary embedding archives with the
auxiliary-vector generation script, results as an \texttt{.rds} object and CSV
tables (including the encoding comparison, congruence matrix, and both loading
matrices), and the regenerated figures. The human response file is not
redistributed where licensing forbids it. The script documents the public-release
download \citep{openpsychometrics-2018} and accepts its location via an
environment variable. Only two steps require Python or network access: the
NLI matrix and the live-embedding demonstrations, which fetch models from the
Hugging Face Hub on first run. All factor analyses, retention methods,
diagnostics, congruence metrics, refinement analyses, and figures are pure R
and run offline. The demonstration was not preregistered.

The reported run used R 4.5.0 \citep{rcoreteam-2025} with semanticfa 0.1.2 \citep{yanitski-westbury-2026}, psych 2.6.5 \citep{revelle-2026}, EGAnet 2.4.1 \citep{golino-christensen-2026}, GPArotation 2026.4-1, Rtsne 0.17, uwot 0.2.4, and reticulate 1.43.0, with seed 42, on an Apple-silicon macOS machine (arm64). Reproducing the demonstration requires semanticfa 0.1.2 or later, whose bundled dataset carries the Qwen3-Embedding-8B vectors analyzed here. The package is on CRAN, with development sources and issue tracker at \url{https://github.com/devon7y/semanticfa}. A getting-started vignette works the bundled inventory end to end.
